Think deep for reef resilience

### 2025-08-10 paper/paper_draft.qmd
Update reef description section

## Reef representations

The spatial extent of the GBR reefs represented in this investigation follows the reef
definitions available via the Great Barrier Reef Marine Park Authority (GBRMPA)
[@GBRMPA2022]. An indicative map of the GBR marine park including the Marine Park management
regions and which reefal bioregion each reef belongs to is shown in [@fig-region-map;
@GBRMPA2024; @GBRMPA2001]. This investigation adopts the GBR reef definition used in Bozec
et al. [-@bozec2025], which includes 3806 reefs across the GBR. Reefs that are not
associated with GBRMPA management areas or reefal bioregions were removed from the
investigation. Additionally, reefs with poor quality bathymetry data coverage or fewer than
three reefs per bioregion were removed from analyses (see Methods - Depth estimation). As a
result, a total of 2429 individual reefs were included in this investigation. Annually and
spatially averaged Degree Heating Week (DHW) data were used to represent future thermal
stress at each reef location. Code used to calculate DHW data and perform analyses involved
in this investigation can be found at **insert repo citation**.

\[**BG: Updated number of reefs and introduced original set of reefs with 3806 to set scene
for justified removal of reefs.**\]

@@

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points). The green polygon represents the Far Northern management area,
the yellow polygon represents the Cairns/Cooktown management area, the pink polygon
represents the Townsville/Whitsunday management area and the blue polygon represents the
Mackay/Capricorn management area [@GBRMPA2024]. The centroid of each reef is represented as
a point for visual clarity, and is coloured according to its GBRMPA reefal bioregion
[@GBRMPA2001]. All 2429 reefs included in this investigation are represented in this figure.
An annotation represents the approximate location of the Australian Institute of Marine
Science Townsville site at Cape
Cleveland.](../outputs/ADRIA_results/HighResCoralStress/figs/region_map.png){#fig-region-map}

>>>>

## Reef definitions

The spatial extent of the GBR reefs represented in this investigation follows the reef
definitions available via the Great Barrier Reef Marine Park Authority (GBRMPA)
[@GBRMPA2022]. An indicative map of the GBR marine park including the Marine Park management
areas [@GBRMPA2024] and the corresponding bioregion for each reef is shown in
[@fig-region-map; @GBRMPA2001]. This investigation adopts the GBR reef outlines as published
by GBRMPA [-@GBRMPA2022], which defines 3806 reefs across the GBR.

Depth characteristics for each reef were extracted using 10m resolution bathymetry data made
available in GBRMPA [-@GBRMPA2020]. For each reef, bathymetry data were spatially matched
with the 3806 reef outlines through geometric intersection analysis. Depth values within
each reef polygon were summarized using zonal statistics to calculate minimum, mean, median,
maximum, and standard deviation of depth values. All depth values were converted to positive
values (depths below sea level). The median depth was used to provide an indicative depth
value for each reef. Code, and the resulting depth estimates, are available at **\[cite
Canonical Reefs repo\]**.

While data adopted for the given assessment are said to be accurate up to 15m from Mean Sea
Level [@GBRMPA2020], there are many sources of uncertainty in the depth estimates used for
this study. These are outlined further in the Discussion. Notably, the adopted depth
estimates were derived from bathymetries based on satellite imagery. Available data can be
limited in their alignment with the established GBRMPA reef outlines used to characterise
reef locations. Quality control flags were assigned to identify potential data limitations:
reefs with no overlapping bathymetry data, reefs with minimum depth values above sea level
(indicating potential data artifacts), and reefs where bathymetry pixels covered less than
5% of the reef polygon area. This removed 842 reefs, or 22%, from the initial set of 3806
reefs. Furthermore, reefs that are not associated with GBRMPA management areas [@GBRMPA2024]
or reefal bioregions [@GBRMPA2001] were also removed from the investigation. Bioregions
which contained fewer than three reefs were also excluded. As a result, 1377 reefs were
excluded (36%), leaving a total of 2429 remaining for the purpose of this investigation.

@@

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points). Colour-coded areas outline GBR marine park management regions,
where green represents the Far Northern management region, yellow represents
Cairns/Cooktown, pink covers Townsville/Whitsunday and blue covers Mackay/Capricorn
[@GBRMPA2024]. Reef centroids are marked individually for visual clarity, and are coloured
according to its bioregion [@GBRMPA2001]. All 2429 reefs included in this investigation are
represented in this figure. An annotation represents the approximate location of the
Australian Institute of Marine Science Townsville site at Cape
Cleveland.](../outputs/ADRIA_results/HighResCoralStress/figs/region_map.png){#fig-region-map
fig-align="center" width="14.8cm"}
